---
always_allow_html: true
---

```{r warning=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
source(here::here("tables", "_common.R"))
```

# Kable vs Flex vs qt {#sec-kableflexgt}

Here I compare a three different ways to make tables. Unfortunately the table numbers and cross-references are still not working as of Quarto version 1.2.335 (Feb 2023), but the developers are aware and a major revamp of cross-referencing is in the works to fix these problems.

```{r}
tbl_cap <- "This should have a number that follows the other tables."
dt <- (mtcars[1:10,])
f1 <- "Here is a footnote about this table"
f2 <- "Here is a second footnote."
```

## Kable

Here is the {kable} table @tbl-kable. Word output is often not good looking. This is a known issue with `kable`.

```{r}
#| label: tbl-kable
#| tbl-cap: !expr paste("kable:", tbl_cap)
#| echo: true
library(knitr)
library(kableExtra)
# note hold_position not working in Quarto v1.0.38.
kbl(dt, booktabs = TRUE) %>%
  kable_styling(latex_options = c("scale_down")) %>%
  kableExtra::footnote(symbol = c(f1, f2))
```

### `kable` quirks

-   Make sure to put `always_allow_html: true` in the yaml at the top of your Rmd or qmd file if outputting to Word. I can't figure out how to put it in the \`
-   Word output is often not good looking. This is a known issue with `kable`
-   Do not pass in caption to `kbl()` if you want to use Quarto's cross-reference engine.
-   `kbl(..., format="pandoc")` can help for Word if your templates stop working but destroys the PDF output.

## `flextable`

Here is the {flextable} table @tbl-flex (note this is broken in Word output). {flextable} gives you a lot more control over your tables with a grammar format (like ggplot2). It also gives nice output to Word, PDF and HTML. Sadly in Quarto v1.2.335 cross-reference and table numbers for Word  is broken, but the developers know about this. The {officer} package, which I think Quarto is leaning on for Word generation, use {flextable} so I am hoping that {flextable} gets moved into the RStudio suite.

```{r}
#| label: tbl-flex
#| tbl-cap: !expr paste("flextable:", tbl_cap)
#| echo: true
library(flextable)
dt %>%
  flextable() %>%
  autofit() %>%
  flextable::footnote(i = 1, j = 1:2,
            value = as_paragraph(c(f1, f2)),
            ref_symbols = c("a", "b"),
            part = "header") %>%
  font(fontname="Times New Roman", part="all") %>%
  fontsize(size=12)
```

### `flextable` quirks

-   Table numbers and cross-refs broken in Word. Known issue.

## `gt`

Here is the {gt} table @tbl-gt (note this is broken in Word output). {gt} has many of the same features as {flextable} and is in the RStudio suite so maybe it'll get more Quarto support in the long run. But like {kable}, the Word output is not good. In addition, currently table captions are broken in Quarto Word output.


```{r}
#| label: tbl-gt
#| tbl-cap: !expr paste("gt:", tbl_cap)
#| echo: true
dt %>%
  gt::gt() %>%
  gt::tab_source_note(source_note = f1) %>%
  gt::tab_source_note(source_note = f2) %>%
  gt::tab_options(
      table.font.size = gt::px(18), 
      table.font.names = "Times New Roman",
      data_row.padding = gt::px(1))
```

### `gt` quirks

- I couldn't get {gt} to respect my font request for HTML unlike for {flextable}
- The footnotes look bad in Word. But I don't know {gt} so probably my error.
- Table captions and numbers broken in Word.

```{=openxml}
<w:p><w:r><w:br w:type="page"/></w:r></w:p>
```
